Analysing Constraint Grammars with a SAT-solver

### 2016-03-09 paper/lrec/eval_futurework.tex
more reporting on the newly found spanish error

\# conflicts found 
 no amb. classes   & 7             & X        & 43      \\ 
{\small (internal~+~interaction)}
                   & {\small
                      (6~+~1)}     & {\small 
                                    (18~+~25)} & {\small (18~+~25)} \\ \hline
% \# false positives 
%  \small{(internal + interaction)}
%                    & 0             & 1 (1 + 0)       & 1 (1 + 0)  \\ \hline

@@

\clock{} no amb. 
           classes & 3 s        & 44 min       & 16 min     \\ \hline

@@

We include two versions of the Spanish grammar in Table~\ref{table:res}: in column \textsc{spa}, we added the lemmas and word forms as described in Section~\ref{sec:realistic_readings}, and in column \textsc{spa}~{\small sep.~lem.}, we just added each word form and lemma as individual readings, allowed to combine with any other reading. 
This latter version ran much faster, but failed to detect an internal conflict for one rule, and reported a false positive for another. The numbers look the same
% FAIL TO DETECT: SELECT:pr_cnjadv_3 cnjadv IF (0 "después de") 
% FALSE POSITIVE: SELECT:este_1 det IF (0 "este" + (m sg)|(m sp)|(mf sg)|(mf sp))

Interestingly, the version $with$ ambiguity classes fails to detect a true interaction conflict, because it---correctly---found an interaction conflict in one of the rules.

As an example of internal conflict, there are two rules that use \texttt{SET Cog = (np cog)}---the problem is that the tag \texttt{cog} does not exist in the Apertium dictionary for Spanish. As another example, four rules require a context word tagged as NP with explicit number, but the lexicon does not indicate any number with NPs. It is likely that this grammar has been written for an earlier version, where such tags have been in place.

>>>>

\# conflicts found 
 w/o~amb. classes   & 7             & X        & 43      \\ 
{\small (internal~+~interaction)}
                   & {\small
                      (6~+~1)}     & {\small 
                                    (18~+~25)} & {\small (18~+~25)} \\ \hline
% \# false positives 
%  \small{(internal + interaction)}
%                    & 0             & 1 (1 + 0)       & 1 (1 + 0)  \\ \hline

@@

\clock{} w/o amb. 
           classes & 3 s        & 44 min       & 16 min     \\ \hline

@@

We include two versions of the Spanish grammar in Table~\ref{table:res}: in column \textsc{spa}, we added the lemmas and word forms as described in Section~\ref{sec:realistic_readings}, and in column \textsc{spa}~{\small sep.~lem.}, we just added each word form and lemma as individual readings, allowed to combine with any other reading. 
This latter version ran much faster, but failed to detect an internal conflict for one rule, and reported a false positive for another. 
% FAIL TO DETECT: SELECT:pr_cnjadv_3 cnjadv IF (0 "después de") 
% FALSE POSITIVE: SELECT:este_1 det IF (0 "este" + (m sg)|(m sp)|(mf sg)|(mf sp))

When we added ambiguity class constraints, we found three more true conflicts, all internal.
Interestingly, the version \emph{with} ambiguity classes fails to detect an interaction conflict, which the simpler version reports, because one of the rules was first detected as internal conflict.

As an example of internal conflict, there are two rules that use \texttt{SET Cog = (np cog)}---the problem is that the tag \texttt{cog} does not exist in the Apertium dictionary for Spanish. As another example, four rules require a context word tagged as NP with explicit number, but the lexicon does not indicate any number with NPs. It is likely that this grammar has been written for an earlier version, where such tags have been in place.
One of the conflicts that was only caught by the ambiguity class constraints had the condition \texttt{IF (1 Comma) (..) (1 CnjCoo)}. The additional constraints correctly prevent the SAT-solver to make comma ambiguous with anything else.
